---
title: "Fama Macbeth"
author: "Sriniwas M"
date: "28 March 2020"
output:
  html_document: default
  pdf_document: default
---
###Update R
```
install.packages("installr")
library(installr)
updateR()
```


###Load the Relevant Packages

```
install.packages("lmtest")
install.packages("rmarkdown")
install.packages("knitr")
install.packages("readxl")
install.packages("dplyr")
install.packages("lme4")

library(lmtest)
library(rmarkdown)
library(knitr)
library(readxl)
library(dplyr)
library(lme4)
```

###Read the Data File and Define Variables

```{r}
setwd("C:/Users/51950005/Desktop/Term III/FFE/fama macbeth assignment")
# ensure that the data has the date column (column 1), followerd by the three FF factors (column 2:4), the risk free rate (column 5), and than the portfolio returns for the 100 portfolios (column 6:105)

Data <- data.frame(read.csv(file = '100_Portfolios_10x10_Daily.csv', header = TRUE)) # read the file. it has first 5 cols date, mkt-rf, smb, hml and rf. next 100 columns as portfolio returns
colnames(Data)[1] <- "Date" # name the first column as Date

# convert to chars to convert to numbers. And than make it a data frame
Data <- lapply(Data, as.character) 
Data <- lapply(Data, as.numeric)
Data <- as.data.frame(Data)

#-----------------------------------------------------------------
#Sanity check of the imported data
#-----------------------------------------------------------------
head(Data)
summary(Data)
names(Data) 

#-----------------------------------------------------------------
# calculate excess returns (i.e. return on ptf - risk free return)
#-----------------------------------------------------------------

for (i in 6:105){
    Data[i] <- Data[i] - Data[5]
}
#------------------------------------------------------------------
#Extract Year from Date information
#------------------------------------------------------------------
nrow(Data)
# 98744 rows i.e. daily observations for each portfolio
# no of years = 2020-1926+1 = 95 years

Data$Date1 <- as.character(Data$Date) 

Data$Date1 <- as.Date(Data$Date1, format="%Y%m%d") # comvert date to year format

Data$year <- as.numeric(format(Data$Date1, "%Y")) # extract year from dates (as number)

#------------------------------------------------------------------
#store portfolio names for reference purposes
#------------------------------------------------------------------
x <- names(Data)
portfolios <- as.vector(x[c(6:105)])

```



### Calculation of beta for portfolios 


**1. Calculation of betas for portfolios using only 1926 year data**
```{r}
Data1926 <- Data[Data$year == 1926,] # create a data frame with only data from year 1926

storage.mktbeta <- rep(NA,100)
storage.sizebeta <- rep(NA,100)
storage.valuebeta <- rep(NA,100)

for (i in 6:105){
    outcome=colnames(Data1926)[i]
    exposure=colnames(Data1926)[2]
    exposuresize=colnames(Data1926)[3]
    exposurebm=colnames(Data1926)[4]
    model <- lm(get(outcome) ~ get(exposure) + get(exposuresize) + get(exposurebm), data=Data1926)
    
    storage.mktbeta[i] <- model$coefficient[2]
    storage.sizebeta[i] <- model$coefficient[3]
    storage.valuebeta[i] <- model$coefficient[4]
  }

storage.mktbeta <- storage.mktbeta[c(6:105)]
storage.sizebeta <- storage.sizebeta[c(6:105)]
storage.valuebeta <- storage.valuebeta[c(6:105)]

# create matrix containing parameters for portfolios in the year 1926

mat1926 <- matrix(c(portfolios, storage.mktbeta, storage.sizebeta, storage.valuebeta),ncol=4) 

```
***Matrix mat1926*** contains the market beta, size factor beta and value factor beta for the 100 portfolios for the year 1926


**2. Calculating betas for portfolios for all years**
```{r}
risk <- list() # a blank list to store the risk matrices (300x3) for each year

for (j in 1926:2020)
  {
  yearlydata <- Data[Data$year == j,]
  storage.mktbeta <- rep(NA,100)
  storage.sizebeta <- rep(NA,100)
  storage.valuebeta <- rep(NA,100)
  for (i in 6:105){
    outcome=colnames(yearlydata)[i]
    exposure=colnames(yearlydata)[2]
    exposuresize=colnames(yearlydata)[3]
    exposurebm=colnames(yearlydata)[4]
    model <- lm(get(outcome) ~ get(exposure) + get(exposuresize) + get(exposurebm), data=yearlydata)
    
    storage.mktbeta[i] <- model$coefficient[2]
    storage.sizebeta[i] <- model$coefficient[3]
    storage.valuebeta[i] <- model$coefficient[4]
  }
  
  storage.mktbeta <- storage.mktbeta[c(6:105)]
  storage.sizebeta <- storage.sizebeta[c(6:105)]
  storage.valuebeta <- storage.valuebeta[c(6:105)]
  risk[[j]] <- matrix(c(portfolios, storage.mktbeta, storage.sizebeta, storage.valuebeta),ncol=4) 
}

# example of risk matrix for year 1926 and year 2020
risk[[1926]]
risk[[2020]]

```
***Risk is a list containing 95 risk matrices*** where each matrix corresponds to a year and has the portfolio betas for the 100 portfolios

**3. NIFTY Bank Return vs Change in M3 Money supply**
```{r}
plot(NIFTYBankReturn ~ M3Change, data=NewData, xlab="Change in M3 money supply", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ M3Change), col="red")
cor.test(NIFTYBankReturn, M3Change, method = "pearson")
```
***No*** correlation between Change in M3 money supply and Bank NIFTY returns

**4. NIFTY Bank Return vs 1 period lagged value of Change in M3 Money supply**
```{r}
plot(NIFTYBankReturn ~ M3ChangeLag, data=NewData, xlab="Lagged change in M3 Money supply", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ M3ChangeLag), col="red")
cor.test(NIFTYBankReturn, M3ChangeLag, method = "pearson")
```
***No*** significant correlation between Lagged change in M3 Money supply and Bank NIFTY returns

**5. NIFTY Bank Return vs DJIA return**
```{r}
plot(NIFTYBankReturn ~ DJIAReturn, data=NewData, xlab="Return on DJIA", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ DJIAReturn), col="red")
cor.test(NIFTYBankReturn, DJIAReturn, method = "pearson")
```
***Positive*** correlation between DJIA returns and Bank NIFTY returns

**6. NIFTY Bank Return vs Change in Fx rate (USD/INR)**
```{r}
plot(NIFTYBankReturn ~ FXRateChange, data=NewData, xlab="Change in USD/INR (INR per USD)", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ FXRateChange), col="red")
cor.test(NIFTYBankReturn, FXRateChange, method = "pearson")
```
***No*** significant correlation between Change in USD/INR rate and Bank NIFTY returns

**7. NIFTY Bank Return vs 1 period lagged Change in Fx rate (USD/INR)**
```{r}
plot(NIFTYBankReturn ~ FXRateChangeLag, data=NewData, xlab="Lagged Change in USD/INR (INR per USD)", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ FXRateChangeLag), col="red")
cor.test(NIFTYBankReturn, FXRateChangeLag, method = "pearson")
```
***Negative*** correlation between Lagged Change in USD/INR rate and Bank NIFTY returns

**8. NIFTY Bank Return vs IIP General Index**
```{r}
plot(NIFTYBankReturn ~ IIPG, data=NewData, xlab="IIP General Index", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ IIPG), col="red")
cor.test(NIFTYBankReturn, IIPG, method = "pearson")
```
***No*** correlation between IIP General Index and Bank NIFTY returns

**9. NIFTY Bank Return vs 1 period lagged IIP General Index**
```{r}
plot(NIFTYBankReturn ~ IIPGLag, data=NewData, xlab="Lagged IIP General Index", ylab="Return on NIFTY Bank")
abline(lm(NIFTYBankReturn ~ IIPGLag), col="red")
cor.test(NIFTYBankReturn, IIPGLag, method = "pearson")
```
***No*** correlation between IIP General Index and Bank NIFTY returns

###Multiple Regression Analysis
**Multiple Regression with All Independent Variables**
```{r}
Fit1 <- lm(NIFTYBankReturn ~ FII + FERChange + M3ChangeLag + DJIAReturn + FXRateChangeLag + IIPGLag, data = NewData)
summary(Fit1)
```
R-square of 49.84%. The independent variables: "Change in FER", "Lagged change in M3 Money supply" and "Lagged valued of IIP General Index" are **insignificant** in determining the Return on NIFTY BANK Index.

**Multiple Regression after Dropping the Insignificant Variables**
```{r}
Fit2 <- lm(NIFTYBankReturn ~ FII + FXRateChangeLag + DJIAReturn, data = NewData)
summary(Fit2)
```
R square of 46.5%.
All the independent variables: "FII", "Lagged change in USD/INR", "and Return on DJIA" are significant in determining the "Return on NIFTY Bank"
FII and Return on DJIA have positive slope coefficients. This implies **Increase in FII and postive returns on DJIA, leads to higher returns on NIFTY Bank.**
But Lagged change in USD/INR has a negative coefficient. This means **when INR depreciates w.r.t. USD, the return on NIFTY Bank decreases.**
